### 2012-01-24 docs/SDC/sdc.tex
tweaks

\section{Implementing an SDC-enabled reaction network (no diffusion)}

\subsection{Introduction to SDC}

@@

\subsection{Strang-split version}

@@

\subsection{SDC version}
In the SDC version, the VODE integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
\{$\rho$, $(\rho h)$, and $(\rho X_k$)\}, including both the advective
fluxes (incorporated via a piecewise constant advective flux source
term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
during the reaction step.  A flowchart of the \maestro\ SDC algorithm is 
shown in figure~\ref{fig:sdc:flowchart}.

@@

\subsubsection{Advective Update}

@@

\subsubsection{Final Update}

@@

At the end of the integration, we define $I_{\rho X_k}$ and $I_{\rho
  h}$.  The actual form of these depends on what we are predicting.
We note that we only need $I_{\rho X_k}$ and $I_{\rho h}$ for the
prediction of the interface states, not the final conservative update.
That is because all we need from the advection solver is the
approximation to $A_{\rho X_k}$ and $A_{\rho h}$, not the final
updated state.

\paragraph{Species Source Terms.}
For the species prediction, the form of $I_{\rho X_k}$ depends on
{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
\begin{itemize}
\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
%
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

@@

\paragraph{Enthalpy Source Terms.}
The obvious analog to the above form for $I_{\rho X_k}$ is to define
\begin{equation}
I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
\label{eq:sdc:Irhoh}
\end{equation}
This can be thought of as an approximation to $(\rho \Hnuc)$,
but this does not match any of the quantities we predict for the
enthalpy update ({\tt enthalpy\_pred\_type}, see \S
\ref{sec:pred:enthalpy}).  The appropriate constructions
are:

@@

\ \\
This is done in {\tt advance.f90} just after the call to {\tt react\_state},
stored in the \multifab\ called {\tt intra}.
These terms are used as the source terms for the
advection step in the next SDC iteration.

\subsection{Summary of Changes}
The major changes from the non-SDC-enabled burners is the addition of
the advective terms to the system of ODEs, the fact that we integrate
$(\rho X_k)$ instead of just $X_k$, and the need to derive the
temperature from the input state for each RHS evaluation by {\tt
  VODE}.

>>>>

\section{SDC Overview}

@@

\section{Strang-split Overview}

@@

\section{SDC Without Thermal Diffusion or Base State Evolution}
In the SDC version, the VODE integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
\{$\rho$, $(\rho h)$, and $(\rho X_k$)\}, including both the advective
fluxes (incorporated via a piecewise constant advective flux source
term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \label{eq:sdc:rhoX} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h} \label{eq:sdc:rhoh}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
during the reaction step.  A flowchart of the \maestro\ SDC algorithm is 
shown in figure~\ref{fig:sdc:flowchart}.

@@

\subsection{Advective Update}

@@

\subsection{Final Update}

@@

Finally we integrate the ODE system (Eqs.~\ref{eq:sdc:rhoX} and \ref{eq:sdc:rhoh}).
At the end of the integration, we define $I_{\rho X_k}$ and $I_{\rho
  h}$.  The actual form of these depends on what we are predicting.
We note that we only need $I_{\rho X_k}$ and $I_{\rho h}$ for the
prediction of the interface states, not the final conservative update.
That is because all we need from the advection solver is the
approximation to $A_{\rho X_k}$ and $A_{\rho h}$, not the final
updated state.

\subsubsection{Species Source Terms.}
For the species prediction, the form of $I_{\rho X_k}$ depends on
{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
\begin{itemize}
\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
%
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

@@

\subsubsection{Enthalpy Source Terms.}
The obvious analog to the above form for $I_{\rho X_k}$ is to define
\begin{equation}
I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
\label{eq:sdc:Irhoh}
\end{equation}
This can be thought of as an approximation to $(\rho \Hnuc)$,
but this does not match any of the quantities we predict for the
enthalpy update ({\tt enthalpy\_pred\_type}, see \S
\ref{sec:pred:enthalpy}).  The appropriate constructions
are:

@@

\subsubsection{Implementation}

This is done in {\tt advance.f90} just after the call to {\tt react\_state},
stored in the \multifab\ called {\tt intra}.
These terms are used as the source terms for the
advection step in the next SDC iteration.

\section{Summary of Changes}
The major changes from the non-SDC-enabled burners is the addition of
the advective terms to the system of ODEs, the fact that we integrate
$(\rho X_k)$ instead of just $X_k$, and the need to derive the
temperature from the input state for each RHS evaluation by {\tt
  VODE}.
